---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  message = FALSE
)
```

# earthquakes

The earthquakes package is created for the Mastering Software Development in R capstone from coursera.

This capstone project will be centered around a dataset obtained from the U.S. National Oceanographic and Atmospheric Administration (NOAA) on significant earthquakes around the world. This dataset contains information about 5,933 earthquakes over an approximately 4,000 year time span.

The overall goal of the capstone project is to integrate the skills you have developed over the courses in this Specialization and to build a software package that can be used to work with the NOAA Significant Earthquakes dataset. This dataset has a substantial amount of information embedded in it that may not be immediately accessible to people without knowledge of the intimate details of the dataset or of R. Your job is to provide the tools for processing and visualizing the data so that others may extract some use out of the information embedded within.

The ultimate goal of the capstone is to build an R package that will contain features and will satisfy a number of requirements that will be laid out in the subsequent Modules. You may want to begin organizing your package and insert various features as you go through the capstone project.

Dataset

* [NOAA Significant Earthquake Database](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1)

## Installation

You can install earthquakes from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("raggaraluz/mastering_r_capstone")
```

## Example

### Downloading data

Data can be downloaded from NOAA Significant Earthquake Database as TSV and converted to a data frame:

```{r, cache = TRUE}
library(readr)

location <- 'https://www.ngdc.noaa.gov/nndc/struts/results?type_0=Exact&query_0=$ID&t=101650&s=13&d=189&dfn=signif.txt'
raw <- read_tsv(location)
```

Then, the dataframe can be cleaned using the earthquakes package functions
```{r}
library(earthquakes)
library(dplyr)
clean <- eq_clean_data(raw) %>%
  eq_location_clean()

knitr::kable(clean %>%
               select(DATE, LATITUDE, LONGITUDE, LOCATION_NAME, EQ_PRIMARY) %>%
               tail)
```

## Timeline geoms

In this package are available two geoms that can be used to plot earthquake data.

The first geom, `geom_timeline()` is used for plotting a time line of earthquakes ranging from xmin to xmax dates, with a point for each earthquake. Optional aesthetics include color, size and alpha. The x aesthetic is a date and an optional y aesthetic is a factor indicating some stratification for multiple lines.

```{r geom_timeline}
library(ggplot2)
library(lubridate)
clean %>%
  filter(COUNTRY %in% c('USA', 'CHINA')) %>%

  ggplot(aes(x = DATE, y = COUNTRY, size=EQ_PRIMARY, fill = DEATHS)) +
    geom_timeline(alpha = 0.25, xmin = ymd('2000-01-01'), xmax = ymd('2015-12-31')) +
    theme_minimal() + theme(panel.grid.minor.x = element_blank(),
                            panel.grid.major.x = element_blank(),
                            axis.line.y = element_blank(),
                            axis.title.y = element_blank(),
                            axis.line.x = element_line(),
                            axis.ticks.x = element_line())
```

